Accurate evaluation of dislocation tractions

### 2021-01-20 src/wip/DanielCelisGarza/Accurate_evaluation_of_dislocation_tractions/paper/main.tex
Corrections to paper.

Specifically, a linear-elastic solid $V$ bounded by a surface $S$ is subjected to traction boundary conditions, $\vec{T}$, on $S_{T}$ and displacement boundary conditions, $\vec{U}$, on $S_{U}$. The ($\tilde{~}$) fields are those generated by the dislocations in an infinite solid and in our case are obtained by evaluating analytic solutions in a DDD simulation \cite{Cai2006}. Formally, the dislocation field satisfies,
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\tilde{\sigma}} = \vec{0}                \\
        \tns{\tilde{\sigma}} = \tns{C} : \tns{\tilde{\varepsilon}} \\
        \tns{\tilde{\varepsilon}} = \frac{1}{2} \left( \nabla \vec{\tilde{u}} + (\nabla \vec{\tilde{u}})^T \right)
    \end{array}
    \right\} \quad \textrm{in V}                                                   \\
    %
     & \tilde{\tns{\sigma}} \cdot \vec{n} = \tilde{\vec{T}} \quad \textrm{on } S_T \\
    %
     & \left.
    \begin{array}{l}
        \vec{\tilde{u}} = \vec{\tilde{u}}, \quad t>0 \\
        \vec{\tilde{u}} = \vec{0}, \quad t=0
        \label{eq:utildebc}
    \end{array}
    \right\} \quad \textrm{on } S_U \,.
\end{align}
%
As the dislocation fields do not vanish on $S$, the dislocations load the volume by generating tractions, $\tilde{\vec{T}}$, on $S_T$ and displacements, $\vec{\tilde{u}}$, on $S_U$. This additional loading deforms $V$, generating an additional ``image'' stress which the dislocations then feel. Therefore corrective ($\hat{~}$) fields must be superimposed to satisfy the desired boundary conditions. The corrective field which accounts for both the applied and image stress is obtained numerically by solving the elastic boundary value problem,
%
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\hat{\sigma}} = \vec{0}              \\
        \tns{\hat{\sigma}} = \tns{C} : \tns{\hat{\varepsilon}} \\
        \tns{\hat{\varepsilon}} = \frac{1}{2} \left( \nabla \vec{\hat{u}} + (\nabla \vec{\hat{u}})^{T} \right)
    \end{array}\right\} \quad \textrm{in V}                                                  \\
    %
     & \hat{\tns{\sigma}}\cdot\vec{n} = \vec{T} - \tilde{\vec{T}} \quad \textrm{on } S_T \label{eq:thatbc} \\
    %
     & \vec{\hat{u}} = \vec{U} - \vec{\tilde{u}} \quad \textrm{on } S_U \,. \label{eq:uhatbc}
\end{align}
%
Once the solutions to both problems are known, their superposition solves the desired mixed boundary value problem,
%
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\sigma} = \nabla \cdot \left( \tns{\hat{\sigma}} + \tns{\tilde{\sigma}} \right) = \vec{0}                                                             \\
        \tns{\sigma} = \tns{\hat{\sigma}} + \tns{\tilde{\sigma}} = \tns{C} : \left( \tns{\hat{\varepsilon}} + \tns{\tilde{\varepsilon}} \right) = \tns{C} : {\tns{\varepsilon}} \\
        \tns{\varepsilon} = \tns{\hat{\varepsilon}} + \tns{\tilde{\varepsilon}} = \frac{1}{2} \left( \nabla ( \vec{\hat{u}} + \vec{\tilde{u}} ) + \left[ \nabla ( \vec{\hat{u}} + \vec{\tilde{u}} ) \right]^T \right) = \frac{1}{2} \left( \nabla \vec{u} + ( \nabla \vec{u} )^{T} \right)
    \end{array}
    \right\}\quad \textrm{in V}                                    \\
    %
     & \tns{\sigma} \cdot \vec{n} = \vec{T} \quad \textrm{on } S_T \\
    %
     & \left.
    \begin{array}{l}
        \vec{u} = \vec{U}, \quad t>0 \\
        \vec{u} = \vec{0}, \quad t=0
    \end{array}
    \right\} \quad \textrm{on } S_U
    \label{eq:ubc}
\end{align}

@@

The original paper by \citet{head1953edge} has a few typos that have been replicated in other sources, and the equations in \cite[p.~59,~64]{hirth1983theory} assume the dislocation is found at the origin. We therefore include the complete and correct expressions in \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew}. \citet{head1953edge} gives two basic cases. \Cref{eq:imageStressAnalyticEdge1} corresponds to the case where $\vec{b}$ is perpendicular to the surface and positive $b$ means it points in the positive $x$-direction,

@@

One serious disadvantage of the analytic tractions is that the implementation of the analytic solution is also much more involved and full of snags. One issue is the calculation of the $y$-coordinate in the local coordinate frame as shown in \cref{f:force_lin_rect} and \cref{eq:problem}. If $\vec{l} \perp \vec{n}$, we get a singularity. As mentioned in \cite{Queyreau}, an easy fix is to rotate the line segment. We do this about its midpoint and around the $\vec{l} \times \vec{n}$ axis in both clockwise and anticlockwise directions. We use the mean of the values as the answer for $\theta = 0$. An example of what this rotation looks like in terms of the forces on a surface element can be seen in \cref{f:rotate} (avoiding the singularity at $\theta = 0$).
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{ftot_rotation_lin_rect_inlaid.pdf}
    \caption{Example of the components of the total force on a surface element (the force summed over all four FE nodes) as a dislocation segment parallel to the surface element ($\theta = 0$) is rotated about its midpoint around the axis defined by $\vec{l}\times\vec{n}$. The force is smooth but not necessarily antisymmetric about the neighbourhood of $\theta=0$.}
    \label{f:rotate}
\end{figure}
The specific shape of the curves will vary depending on the element-segment configuration, but is smooth and well-behaved about singularity. For our purposes, we use a total of 8 perturbations of $1\deg$ each (4 clockwise and 4 anti-clockwise). We found this worked well, but can be changed if desired.

@@

The resulting force vector is then used in \cref{eq:thatbc} to calculate $\tns{\hat{\sigma}}$, since $\tilde{\vec{T}} \equiv \tilde{\vec{F}}$. \Cref{f:headvstractionfem} shows the simple system we used to compare the image stresses calculated by our FE solver using numeric tractions v.s. analytic tractions v.s. infinite-domain, singular image stresses in \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew}. The three test cases are: two edge dislocations and one screw, all of which have line direction $\vec{l} = [0\, 0\, 1]$. The Burgers vectors for the different scenarios are $\vec{b}_\textrm{e1} \equiv \vec{b} = [1\, 0\, 0]$, $\vec{b}_\textrm{e2} \equiv \vec{b} = [0\, 1\, 0]$, as defined in \cite{head1953edge}; and $\vec{b}_\textrm{s} \equiv \vec{b} = \vec{l} = [0\, 0\, 1] $, as defined in \cite[p.~59,~64]{hirth1983theory}.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{head_vs_fe_tractions_setup_diagram.pdf}
    \caption{Dislocation parallel to a surface described by the line $x = 0$, where the $(x,\,y)$ dislocation coordinates are $(a,\,c)$ and the dislocation line going in the positive $z$-direction. (a) describes the box we used for our comparison, a  $40 \times 40 \times 40$ element cubic box with side lengths equal to $2000\, \lVert \vec{b} \rVert$ with $\vec{T} = \vec{0}$, traction boundary conditions only on the nodes along $x = 0$, and $\vec{U} = \vec{0}$, displacement conditions everywhere else. (b) is the 2D view with the dislocation coordinates $(a,~c)$, as well as the two edge Burgers vectors in \cite{head1953edge}, $\vec{b}_\textrm{e1} \equiv \vec{b} = [1\, 0\, 0]$, $\vec{b}_\textrm{e2} \equiv \vec{b} = [0\, 1\, 0]$ and the screw Burgers vector $\vec{b}_\textrm{s} \equiv \vec{b} = \vec{l} = [0\, 0\, 1] $ in \cite[p.~59,~64]{hirth1983theory}.}
    \label{f:headvstractionfem}
\end{figure}
The slices we took for our contour plots in \cref{s:results} are on the middle plane of the domain at $z = 1000\, $.

@@

However, we found a simple and realistic configuration that is close to the setup in the infinite domain examples. The differences between those comparisons and the simulation are described in \cref{t:simulation_params}. We use a mobility law developed in-house by B. Bromage \cite{bromage2018calculating}, that corrects common issues found in other laws. We also rotate the domain such that the $[1\, 1\, 1]$ and $[1\, \overline{1}\, 0]$ crystallographic directions respectively correspond to the simulation's $x$ and $y$-directions, for this we use the same rotation technique as \cite{YU2018}. The dislocation was allowed to move under no external loads or displacements, with $\vec{T} = \vec{0}$ boundary conditions only on the $yz$-plane and $\vec{U} = \vec{0}$ everywhere else.
\begin{table}
    \centering
    \caption{Parameters for our simulation. Where $\mtx{R}$ is a rotation matrix such that the $\vec{x},\, \vec{y},\, \vec{z}$ basis vectors of our simulation correspond to the $\vec{b},\, \vec{n},\, \vec{l}$ crystallographic directions, i.e. $\mtx{R} \vec{x} = \vec{b}$. Everything else was kept just as the previous comparisons.}
    \label{t:simulation_params}
    \begin{tabular}{ll}
        \toprule
        Parameter          & Value                                           \\
        \midrule
        Crystal Structure  & BCC                                             \\
        $\vec{b}$          & $[1\, 1\, 1]$                                   \\
        $\vec{n}$          & $[1\, \overline{1}\, 0]$                        \\
        $\vec{l}$          & $[1\, 1\, \overline{2}]$                        \\
        $b$                & $\sqrt{3}/2$                                    \\
        $a$                & $5\, $                                          \\
        Grid Size          & $(20,\, 20,\, 20)$                              \\
        Domain Size        & $(2000,\, 2000,\, 2000)\, $                     \\
        Lattice size       & $3.18\times 10^{-4}~\mu\text{m}$                \\
        $(x_{0},\, y_{0})$ & $(62.5,\, 1000)\, $                             \\
        Min segment length & $50\, $                                         \\
        Max segment length & $125\, $                                        \\
        $\mtx{R}$          & $[\hat{\vec{b}}~ \hat{\vec{n}}~ \hat{\vec{l}}]$ \\
        \bottomrule
    \end{tabular}
\end{table}

@@

\Cref{f:rel_err_perp_edge} shows that even for dislocations only one dislocation core radius ($5\, b$) away from the surface element, the force can be obtained, up to numerical precision, with 1000 Gauss quadrature points $Q$ for all segment lengths tested. It also shows a very peculiar issue Gauss quadrature has when computing integrals of rational functions when the Gauss nodes are close poles/maximal values. This undesirable behaviour is observed in the case where $Q = 11$. Where the highest weighted Gauss node is closest to the point where $1/R_{a}$ is maximal, resulting in lower accuracy when compared to $Q = 2, 10$ in \cref{f:rel_err_perp_edge} (a) and (b).
\begin{figure}
    \centering
    \subfloat[Segment starts a single dislocation core radius away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.475\linewidth]{perp_e_xz=5.pdf}}
    }
    ~
    \subfloat[Segment starts 3 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.475\linewidth]{perp_e_xz=15_leg.pdf}}
    }

    \subfloat[Segment starts 201 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.475\linewidth]{perp_e_xz=1005_leg.pdf}}
    }
    ~
    \subfloat[Segment starts 20001 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.47\linewidth]{perp_e_xz=100005_leg.pdf}}
    }
    \caption{Log-log plot of the relative error as a function of dislocation segment length for a perpendicular edge dislocation (\cref{f:gauss_quad_test}a). $x^{1}_{z}$ is the $z$-coordinate of node $\mathbf{x_1}$. $\vec{b} = [1\, 0\, 0]$, line direction, $\vec{l} = [0\, 0\, 1]$, and dislocation core radius, $a = 5\, $, the surface element's normal and size are, $\vec{n} = [0\, 0\, 1]$, $L = 1000\, $, respectively. $Q$ is the number of quadrature points per dimension.}
    \label{f:rel_err_perp_edge}
\end{figure}
It is worth noting however that the relative errors for small numbers of quadrature points don't really start decreasing until relatively large distances. And when close to a surface, these can be quite large even under hihgly symmetric circumstances.

The limitations become even more evident when the dislocation line segment is parallel to a surface element. In \cref{f:rel_err_par_edge}, we observe the relative errors are quite large when close to the surface. At distances larger than $10^4\, $, loss of significance causes the relative errors to converge at $\sim 1$ which is expected as two finite precision floating point numbers get closer to zero. Of particular note is how large the relative errors are even at $100\, $ away from the surface, even for large numbers of quadrature points. This figure backs up the earlier point regarding Gauss nodes close to maximal values of rational functions. It can be shocking to see that as $Q = 2$ performs significantly better than $Q = 10, 11, 100, 100$ at distances from the surface as having nodes near the maximal values is a large source of error. Consequently, different configurations have different optimal numbers of points. The numerical instability of this method, which when coupled to the chaotic nature of dislocation dynamics and stiffness of the equations of motion, can lead to large deviations between simulations.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{par_e_z.pdf}
    \caption{Log-log plot of the relative error as a function of distance from the surface element for a parallel edge dislocation (\cref{f:gauss_quad_test}b). Burgers vector, $\vec{b} = [0\, 1\, 0]$, line direction, $\vec{l} = [1\, 0\, 0]$, dislocation core and surface element parameters are the same as \cref{f:rel_err_perp_edge}. The dislocation length is fixed to $10^{6}\, b,\, x\in\left[-0.5\times10^{6},\, 0.5\times10^{6} \right]$ and bisects the surface element along the $[1\, 0\, 0]$ direction. The whole dislocation was segmented into $10^4$ pieces of length $100\, $ to prevent the dislocation from intersecting the surface element when they were rotated to avoid the singularity. The relative error at large distances ($>10^6\, $) converges to one due to loss of significance.}
    \label{f:rel_err_par_edge}
\end{figure}

@@

    \subfloat[\label{sf:xzepar}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{sxzparEdgeFiveb.pdf}}
    }
    ~
    \subfloat[\label{sf:yzepar}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{syzparEdgeFiveb.pdf}}
    }
    ~
    \subfloat[\label{sf:zzepar}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{szzparEdgeFiveb.pdf}}
    }
    \caption{(a) to (c) show the relevant stress fields from a dislocation in the same configuration that yields the plots found in \cref{f:rel_err_perp_edge} as shown in \cref{f:gauss_quad_test}(a), where the closest node is a dislocation core radius away from the surface, $a = 5\, $. (e) to (f) does the same for the configuration that yields \cref{f:rel_err_par_edge} as shown in \cref{f:gauss_quad_test}(b), where the whole dislocation is a core radius away from the surface, $a = 5\, $. The white line or dot is the dislocation.}
    \label{f:sigma_edge_test_config}
\end{figure}

@@

    \caption{Image stresses for an edge dislocation with $\vec{l} = [0 0 1]$, $\vec{b} = [1 0 0]$, where $a = 5\, $, with coordinates $(25.6410,\, 1000)\, $, i.e. the centre of the first FE from the surface at $x=0$, and in the centre of the simulation box from top to bottom. From left to right, the subfigure columns correspond to the stress fields for the infinite-domain solution (no superscript), analytic tractions + FEM (A superscript), and numeric tractions ($Q = 1$) + FEM (N superscript). The subfigure rows from top to bottom represent $\hat{\sigma}_{xx}$, $\hat{\sigma}_{yy}$ and $\hat{\sigma}_{xy}$.}
    \label{f:head_vs_ana_vs_num_eperp}
\end{figure}
As stated in \cref{s:method}, tractions are used to calculate the image stresses resulting from the boundary conditions. We therefore compare the differences in image stresses resulting from numeric ($Q = 1$) and analytic traction calculations of both our implementations and how they compare to the infinite-domain, singular expressions in \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew}\footnote{Since the first term in each equation corresponds to the real stresses, we omit them to view the image stresses.}. The stress field comparisons for all three cases are found in \cref{f:head_vs_ana_vs_num_eperp,f:head_vs_ana_vs_num_epar,f:head_vs_ana_vs_num_screw}, where the dislocation is denoted by a white dot.

>>>>

Specifically, a linear-elastic solid $V$ bounded by a surface $S$ is subjected to traction boundary conditions, $\vec{T}$, on $S_{T}$ and displacement boundary conditions, $\vec{U}$, on $S_{U}$. The ($\tilde{~}$) fields are those generated by the dislocations in an infinite solid and in our case are obtained by evaluating analytic solutions in a DDD simulation \cite{Cai2006}. Formally, the dislocation field satisfies,
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\tilde{\sigma}} = \vec{0}                \\
        \tns{\tilde{\sigma}} = \tns{C} : \tns{\tilde{\varepsilon}} \\
        \tns{\tilde{\varepsilon}} = \frac{1}{2} \left( \nabla \vec{\tilde{u}} + (\nabla \vec{\tilde{u}})^T \right)
    \end{array}
    \right\} \quad \textrm{in V}                                                   \\
    %
     & \tilde{\tns{\sigma}} \cdot \vec{n} = \tilde{\vec{T}} \quad \textrm{on } S_T \\
    %
     & \left.
    \begin{array}{l}
        \vec{\tilde{u}} = U, \quad t>0 \\
        \vec{\tilde{u}} = \vec{0}, \quad t=0
        \label{eq:utildebc}
    \end{array}
    \right\} \quad \textrm{on } S_U \,.
\end{align}
%
As the dislocation fields do not vanish on $S$, the dislocations load the volume by generating tractions, $\tilde{\vec{T}}$, on $S_T$ and displacements, $\vec{\tilde{u}}$, on $S_U$. This additional loading deforms $V$, generating an additional ``image'' stress which the dislocations then feel. Therefore corrective ($\hat{~}$) fields must be superimposed to satisfy the desired boundary conditions. The corrective field which accounts for both the applied and image stress is obtained numerically by solving the elastic boundary value problem,
%
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\hat{\sigma}} = \vec{0}              \\
        \tns{\hat{\sigma}} = \tns{C} : \tns{\hat{\varepsilon}} \\
        \tns{\hat{\varepsilon}} = \frac{1}{2} \left( \nabla \vec{\hat{u}} + (\nabla \vec{\hat{u}})^{T} \right)
    \end{array}\right\} \quad \textrm{in V}                                                  \\
    %
     & \hat{\tns{\sigma}}\cdot\vec{n} = \vec{T} - \tilde{\vec{T}} \quad \textrm{on } S_T \label{eq:thatbc} \\
    %
     & \vec{\hat{u}} = \vec{U} - \vec{\tilde{u}} \quad \textrm{on } S_U \,. \label{eq:uhatbc}
\end{align}
%
Once the solutions to both problems are known, their superposition solves the desired mixed boundary value problem,
%
\begin{align}
     & \left.
    \begin{array}{l}
        \nabla \cdot \tns{\sigma} = \nabla \cdot \left( \tns{\hat{\sigma}} + \tns{\tilde{\sigma}} \right) = \vec{0}                                                             \\
        \tns{\sigma} = \tns{\hat{\sigma}} + \tns{\tilde{\sigma}} = \tns{C} : \left( \tns{\hat{\varepsilon}} + \tns{\tilde{\varepsilon}} \right) = \tns{C} : {\tns{\varepsilon}} \\
        \tns{\varepsilon} = \tns{\hat{\varepsilon}} + \tns{\tilde{\varepsilon}} = \frac{1}{2} \left( \nabla ( \vec{\hat{u}} + \vec{\tilde{u}} ) + \left[ \nabla ( \vec{\hat{u}} + \vec{\tilde{u}} ) \right]^T \right) = \frac{1}{2} \left( \nabla \vec{u} + ( \nabla \vec{u} )^{T} \right)
    \end{array}
    \right\}\quad \textrm{in V}                                    \\
    %
     & \tns{\sigma} \cdot \vec{n} = \vec{T} \quad \textrm{on } S_T \\
    %
     & \left.
    \begin{array}{l}
        \vec{u} = \vec{U}, \quad t>0 \\
        \vec{u} = \vec{0}, \quad t=0
    \end{array}
    \right\} \quad \textrm{on } S_U
    \label{eq:ubc}
\end{align}

@@

The original paper by \citet{head1953edge} has a few typos that have been replicated in other sources. We therefore include the complete and correct expressions in \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew}. \citet{head1953edge} gives two basic cases. \Cref{eq:imageStressAnalyticEdge1} corresponds to the case where $\vec{b}$ is perpendicular to the surface and positive $b$ means it points in the positive $x$-direction,

@@

One serious disadvantage of the analytic tractions is that the implementation of the analytic solution is also much more involved and full of snags. One issue is the calculation of the $y$-coordinate in the local coordinate frame as shown in \cref{f:force_lin_rect} and \cref{eq:problem}. If $\vec{l} \perp \vec{n}$, we get a singularity. As mentioned in \cite{Queyreau}, an easy fix is to rotate the line segment. We do this about its midpoint and around the $\vec{l} \times \vec{n}$ axis in both clockwise and anticlockwise directions. We use the mean of the values as the answer for $\theta = 0$. An example of what this rotation looks like in terms of the forces on a surface element can be seen in \cref{f:rotate} (avoiding the singularity at $\theta = 0$).
\begin{figure}
    \centering
    \subfloat[]{\includegraphics[width=0.45\linewidth]{ftot_rotation_lin_rect.pdf}}
    ~
    \subfloat[]{\includegraphics[width=0.45\linewidth]{ftot_rotation_lin_rect_zoom.pdf}}
    % \includegraphics[width=0.8\linewidth]{ftot_rotation_lin_rect_inlaid.pdf}
    \caption{(a) Example of the components of the total force on a surface element (the force summed over all four FE nodes) as a dislocation segment parallel to the surface element ($\theta = 0$) is rotated about its midpoint around the axis defined by $\vec{l}\times\vec{n}$. (b) is zomms into $\pm 10 \deg$, the force is smooth but not necessarily antisymmetric about the neighbourhood of $\theta=0$.}
    \label{f:rotate}
\end{figure}
The specific shape of the curves will vary depending on the element-segment configuration, but is smooth and well-behaved about singularity. For our purposes, we use a total of 8 perturbations of $1\deg$ each (4 clockwise and 4 anti-clockwise). We found this worked well, but can be changed if desired.

@@

The resulting force vector is then used in \cref{eq:thatbc} to calculate $\tns{\hat{\sigma}}$, since $\tilde{\vec{T}} \equiv \tilde{\vec{F}}$. \Cref{f:headvstractionfem} shows the simple system we used to compare the image stresses calculated by our FE solver using numeric tractions v.s. analytic tractions v.s. infinite-domain, singular image stresses in \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew}. The three test cases are: two edge dislocations and one screw, all of which have line direction $\vec{l} = [0\, 0\, 1]$. The Burgers vectors for the different scenarios are $\vec{b}_\textrm{e1} \equiv \vec{b} = [1\, 0\, 0]$, $\vec{b}_\textrm{e2} \equiv \vec{b} = [0\, 1\, 0]$, as defined in \cite{head1953edge}; and $\vec{b}_\textrm{s} \equiv \vec{b} = \vec{l} = [0\, 0\, 1] $, as defined in \cite[p.~59,~64]{hirth1983theory}.

The units in all our examples are normalised to lattice parameter, $a$ is the dislocation core radius for the non-singular formulation discussed in \cite{Cai2006}, and $b \equiv \lVert \vec{b} \rVert$.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{head_vs_fe_tractions_setup_diagram.pdf}
    \caption{Dislocation parallel to a surface described by the line $x = 0$, where the $(x,\,y)$ dislocation coordinates are $(a,\,c)$ and the dislocation line going in the positive $z$-direction. (a) describes the box we used for our comparison, a  $40 \times 40 \times 40$ element cubic box with side lengths equal to $2000$ with $\vec{T} = \vec{0}$, traction boundary conditions only on the nodes along $x = 0$, and $\vec{U} = \vec{0}$, displacement conditions everywhere else. (b) is the 2D view with the dislocation coordinates $(a,~c)$, as well as the two edge Burgers vectors in \cite{head1953edge}, $\vec{b}_\textrm{e1} \equiv \vec{b} = [1\, 0\, 0]$, $\vec{b}_\textrm{e2} \equiv \vec{b} = [0\, 1\, 0]$ and the screw Burgers vector $\vec{b}_\textrm{s} \equiv \vec{b} = \vec{l} = [0\, 0\, 1] $ in \cite[p.~59,~64]{hirth1983theory}.}
    \label{f:headvstractionfem}
\end{figure}
The slices we took for our contour plots in \cref{s:results} are on the middle plane of the domain at $z = 1000\, $.

@@

However, we found a simple and realistic configuration that is close to the setup in the infinite domain examples. The differences between those comparisons and the simulation are described in \cref{t:simulation_params}. We use a mobility law developed in-house by B. Bromage \cite{bromage2018calculating}, that corrects common issues found in other laws. We also rotate the domain such that the $[1\, 1\, 1]$ and $[1\, \overline{1}\, 0]$ crystallographic directions respectively correspond to the simulation's $x$ and $y$-directions, for this we use the same rotation technique as \cite{YU2018}. The dislocation was allowed to move under no external loads or displacements, with $\vec{T} = \vec{0}$ boundary conditions only on the $yz$-plane and $\vec{U} = \vec{0}$ everywhere else.
\begin{table}
    \centering
    \caption{Parameters for our simulation. Where $\mtx{R}$ is a rotation matrix such that the $\vec{x},\, \vec{y},\, \vec{z}$ basis vectors of our simulation correspond to the $\vec{b},\, \vec{n},\, \vec{l}$ crystallographic directions, i.e. $\mtx{R} \vec{x} = \vec{b}$. Everything else was kept just as the previous comparisons. All values are in units of lattice parameters.}
    \label{t:simulation_params}
    \begin{tabular}{ll}
        \toprule
        Parameter          & Value                                           \\
        \midrule
        Crystal Structure  & BCC                                             \\
        $\vec{b}$          & $[1\, 1\, 1]$                                   \\
        $\vec{n}$          & $[1\, \overline{1}\, 0]$                        \\
        $\vec{l}$          & $[1\, 1\, \overline{2}]$                        \\
        $b$                & $\sqrt{3}/2$                                    \\
        $a$                & $5\, $                                          \\
        Grid Size          & $(20,\, 20,\, 20)$                              \\
        Domain Size        & $(2000,\, 2000,\, 2000)\, $                     \\
        Lattice size       & $3.18\times 10^{-4}~\mu\text{m}$                \\
        $(x_{0},\, y_{0})$ & $(62.5,\, 1000)\, $                             \\
        Min segment length & $50\, $                                         \\
        Max segment length & $125\, $                                        \\
        $\mtx{R}$          & $[\hat{\vec{b}}~ \hat{\vec{n}}~ \hat{\vec{l}}]$ \\
        \bottomrule
    \end{tabular}
\end{table}

@@

\Cref{f:rel_err_perp_edge} shows that even for dislocations only one dislocation core radius ($5$) away from the surface element, the force can be obtained, up to numerical precision, with 1000 Gauss quadrature points $Q$ for all segment lengths tested. It also shows a very peculiar issue Gauss quadrature has when computing integrals of rational functions when the Gauss nodes are close poles/maximal values. This undesirable behaviour is observed in the case where $Q = 11$. Where the highest weighted Gauss node is closest to the point where $1/R_{a}$ is maximal, resulting in lower accuracy when compared to $Q = 2, 10$ in \cref{f:rel_err_perp_edge} (a) and (b).
\begin{figure}
    \centering
    \subfloat[Segment starts a single dislocation core radius away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.475\linewidth]{perp_e_xz=5.pdf}}
    }
    ~
    \subfloat[Segment starts 3 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.475\linewidth]{perp_e_xz=15_leg.pdf}}
    }

    \subfloat[Segment starts 201 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.475\linewidth]{perp_e_xz=1005_leg.pdf}}
    }
    ~
    \subfloat[Segment starts 20001 dislocation core radii away.]
    {
        {\includegraphics[trim={0.5cm 0.55cm 0.65cm 0.1cm},clip,width=0.47\linewidth]{perp_e_xz=100005_leg.pdf}}
    }
    \caption{Log-log plot of the relative error as a function of dislocation segment length for a perpendicular edge dislocation (\cref{f:gauss_quad_test}a). $x^{1}_{z}$ is the $z$-coordinate of node $\mathbf{x_1}$. $\vec{b} = [1\, 0\, 0]$, line direction, $\vec{l} = [0\, 0\, 1]$, and dislocation core radius, $a = 5$, the surface element's normal and size are, $\vec{n} = [0\, 0\, 1]$, $L = 1000$, respectively. $Q$ is the number of quadrature points per dimension.}
    \label{f:rel_err_perp_edge}
\end{figure}
It is worth noting however that the relative errors for small numbers of quadrature points don't really start decreasing until relatively large distances. And when close to a surface, these can be quite large even under hihgly symmetric circumstances.

The limitations become even more evident when the dislocation line segment is parallel to a surface element. In \cref{f:rel_err_par_edge}, we observe the relative errors are quite large when close to the surface. At distances larger than $10^4\, $, loss of significance causes the relative errors to converge at $\sim 1$ which is expected as two finite precision floating point numbers get closer to zero. Of particular note is how large the relative errors are even at $100$ lattice units away from the surface, even for large numbers of quadrature points. This figure backs up the earlier point regarding Gauss nodes close to maximal values of rational functions. It can be shocking to see that as $Q = 2$ performs significantly better than $Q = 10, 11, 100, 100$ at distances from the surface as having nodes near the maximal values is a large source of error. Consequently, different configurations have different optimal numbers of points. The numerical instability of this method, which when coupled to the chaotic nature of dislocation dynamics and stiffness of the equations of motion, can lead to large deviations between simulations.
\begin{figure}
    \centering
    \includegraphics[width=0.8\linewidth]{par_e_z.pdf}
    \caption{Log-log plot of the relative error as a function of distance from the surface element for a parallel edge dislocation (\cref{f:gauss_quad_test}b). Burgers vector, $\vec{b} = [0\, 1\, 0]$, line direction, $\vec{l} = [1\, 0\, 0]$, dislocation core and surface element parameters are the same as \cref{f:rel_err_perp_edge}. The dislocation length is fixed to $10^{6},\, x\in\left[-0.5\times10^{6},\, 0.5\times10^{6} \right]$ and bisects the surface element along the $[1\, 0\, 0]$ direction. The whole dislocation was segmented into $10^4$ pieces of length $100$ to prevent the dislocation from intersecting the surface element when they were rotated to avoid the singularity. The relative error at large distances ($>10^6\, $) converges to one due to loss of significance.}
    \label{f:rel_err_par_edge}
\end{figure}

@@

    \subfloat[\label{sf:xzepar}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{sxzparEdgeFiveb.pdf}}
    }
    ~
    \subfloat[\label{sf:yzepar}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{syzparEdgeFiveb.pdf}}
    }
    ~
    \subfloat[\label{sf:zzepar}]
    {
        {\includegraphics[trim={0.5cm 0cm 0cm 0cm},clip,width=0.3\linewidth]{szzparEdgeFiveb.pdf}}
    }
    \caption{(a) to (c) show the real stress fields from a dislocation in the same configuration that yields the plots found in \cref{f:rel_err_perp_edge} as shown in \cref{f:gauss_quad_test}(a), where the closest node is a dislocation core radius away from the surface, $a = 5$. (e) to (f) does the same for the configuration that yields \cref{f:rel_err_par_edge} as shown in \cref{f:gauss_quad_test}(b), where the whole dislocation is a core radius away from the surface, $a = 5$. The white line or dot is the dislocation. Units are in terms of lattice parameters.}
    \label{f:sigma_edge_test_config}
\end{figure}

@@

    \caption{Image stresses for an edge dislocation with $\vec{l} = [0 0 1]$, $\vec{b} = [1 0 0]$, where $a = 5$, with coordinates $(26,\, 1000)$, i.e. the centre of the first FE from the surface at $x=0$, and in the centre of the simulation box along the $y$-direction. (a), (e) and (i) are the stress fields for the infinite-domain solution, $\tns{\hat{\sigma}}$; (b), (f) and (j) are those obtaine from analytic tractions + FEM, $\tns{\hat{\sigma}}^{\textrm{A}}$; (c), (g) and (k) are those obtained using numeric tractions ($Q = 1$) + FEM, $\tns{\hat{\sigma}}^{\textrm{N}}$. (a) to (c) represent the $xx$; (e) to (g) the $yy$; and (i) to (k) the $xy$ components of the stress tensor.}
    \label{f:head_vs_ana_vs_num_eperp}
\end{figure}
As stated in \cref{s:method}, tractions are used to calculate the image stresses resulting from the boundary conditions. We therefore compare the differences in image stresses resulting from numeric ($Q = 1$) and analytic traction calculations of both our implementations and how they compare to the infinite-domain, singular expressions in \cref{eq:imageStressAnalyticEdge1,eq:imageStressAnalyticEdge2,eq:imageStressAnalyticScrew}\footnote{Since the first term in each equation corresponds to the real stresses, we omit them to view the image stresses.}. The stress field comparisons for all three cases are found in \cref{f:head_vs_ana_vs_num_eperp,f:head_vs_ana_vs_num_epar,f:head_vs_ana_vs_num_screw}, where the dislocation is denoted by a white dot.
